\section{Conclusion and Limitations}
\label{sec:conclusion}

We presented \PTA{} as a minimal test of active in-episode context acquisition under hidden dynamics. The core point is that probing is an intervention: it gives the network a response trace, but it also changes the physical state that the task policy must control. In the cross-material Genesis benchmark, this mechanism improves transfer on the elastoplastic OOD split but underperforms a matched reactive baseline on the other four splits. The result is intentionally asymmetric: the paper reports a conditional finding about when probing is physically worth doing, not a broad robustness claim.

The pattern is consistent with, but does not validate, a recoverable-response interpretation. Probing appears useful when the probe-induced perturbation relaxes on the task timescale, as in the tested elastoplastic split, and detrimental when it irreversibly rearranges the substrate, as in the granular splits. In this reading, the same learned context mechanism succeeds or fails depending on whether material physics mitigates the distribution shift created to obtain that context. The M8 diagnostic is narrower: this single-seed sand-only privileged-parameter baseline did not extrapolate to elastoplastic. The recoverability account remains post-hoc and under-constrained: the evidence comes from one positive material family, three seeds with high variance, and a simulator-specific setup.

Several limitations define the next steps. This is a simulator-state diagnostic benchmark: the probe is scripted rather than optimized for information gain; the observation includes simulator-derived aggregate-particle statistics; all training is sand-only; and no sim-to-real transfer is evaluated. Future work should learn probe policies and probe-gating decisions, test additional viscoelastic and granular media, characterize the learned latent representation, and evaluate probe-conditioned policies with realistic tactile or visual estimators.
