Hidden physical dynamics make zero-shot control difficult: visually similar material states can respond differently to contact. We study when active probing helps, using \PTA{}, a deliberately minimal two-stage framework that executes a fixed probe, encodes the resulting proprioceptive and simulator-derived aggregate-particle trace into a latent task-conditioning vector, and conditions a residual policy on that vector. We evaluate on a Genesis edge-push diagnostic benchmark spanning sand, snow, and elastoplastic media; all methods train on sand only and transfer zero-shot. The empirical picture is asymmetric. On OOD elastoplastic, \PTA{} improves transfer efficiency by 14.7 percentage points (60.7\% vs.\ 46.0\%) and reduces spillage by 14.7 percentage points (39.3\% vs.\ 54.0\%) over reactive PPO; ablations suggest that both the fixed probe and latent conditioning contribute to this split-specific gain. On ID sand, two sand parameter shifts, and snow, \PTA{} underperforms by 13--22 percentage points. We therefore do not claim broad cross-material robustness. The pattern is consistent with, but does not validate, a recoverable-response interpretation: probing helped in the tested elastoplastic split and hurt in the tested granular or partly irreversible splits. A single-seed privileged-parameter diagnostic scores 0\% transfer on elastoplastic; we read this narrowly as a failure of this sand-only parameter-conditioned baseline, not as a general comparison between passive parameters and interaction traces. \PTA{} is best read as a diagnostic tool for studying when active context acquisition helps or hurts under hidden dynamics.
